gij bent 2

### 2025-08-22 paper/paper.qmd
references + place in the S-curve

Within the propagation of linguistic change, different roles can be distinguished. Innovators, who are hypothesized to be less central and only loosely connected to others through weak ties, acquiring novel ideas from outside their immediate community [@rogers_diffusion_2003: 282-283]. Early adopters, by contrast, are hypothesized to be more centrally positioned and to enjoy a higher degree of connectivity within the population. Through their many strong ties, they amplify and legitimize the innovation, spreading it efficiently across their dense local networks [@rogers_diffusion_2003: 282-283]. In this way, innovators and early adopters together bridge the gap between the periphery and the core of a community, ensuring that new linguistic forms do not remain isolated but instead gain broader social traction [@granovetter_strength_1973;@milroy_linguistic_1985]. The idea that innovation originates in the periphery of a network before it can spread has been attested before in network studies [@van_de_leemput_transformation_2024], but not whithin linguistics.

@@

>>>>

Within the propagation of innovations, different roles can be distinguished. Innovators, who are less central and only loosely connected to others through weak ties, acquiring novel ideas from outside their immediate community [@rogers_diffusion_2003: 282-283]. Early adopters, by contrast, are more centrally positioned and enjoy a higher degree of connectivity within the population. Through their many strong ties, they amplify and legitimize the innovation, spreading it efficiently across their dense local networks [@rogers_diffusion_2003: 282-283]. In this way, innovators and early adopters together bridge the gap between the periphery and the core of a community, ensuring that new linguistic forms do not remain isolated but instead gain broader social traction [@granovetter_strength_1973;@milroy_linguistic_1985]. The idea that innovation originates in the periphery of a network before it can spread has been attested before in network studies [@van_de_leemput_transformation_2024], but not whithin linguistics.

@@

The observed synchronic variation can also be interpreted through the framework of binary language change, where a new variant, *gij bent*, competes with an older one, *gij zijt*. Such binary changes are typically modeled using an S-curve [@speelman_logistic_2014], which plots the cumulative adoption of the innovation over time. This S-curve has a bell-shaped counterpart, which does not capture the cumulative adoption, but instead captures the rate of adoption at any given point in time: few individuals adopt at the very beginning, adoption peaks as the innovation gains momentum, and then levels off once most speakers have shifted. Diffusion research has shown that the individuals driving this process can be categorized into five adopter groups, defined mathematically by standard deviations around the mean adoption time [@rogers_diffusion_2003: PAGES]: innovators, early adopters, early majority, late majority, and laggards (see FIGURE BELOW). As already hinted at in @sec-introduction, each of these categories is embedded in the network in a different way. Innovators, positioned at the periphery and connected through weak ties, acquire innovations from outside their immediate community and initiate change. Early adopters, in contrast, are more centrally embedded and exert strong influence through their many strong ties, thereby accelerating diffusion and granting legitimacy to the innovation [REFS]. Our finding that more influential Twitter users are more likely to employ *gij bent* suggests that this form has moved beyond the initial innovator stage and is now being amplified by early adopters. In this way, we can infer in which stage of the S-curve the change is situated: between the second and the first standard deviation before the mean.

```{r adopter-categories}
# Load libraries
library(ggplot2)

# Simulated time steps
time <- seq(0, 20, 0.1)

# Logistic S-curve (cumulative adoption)
s_curve <- 100 / (1 + exp(-0.6 * (time - 10)))

# Bell curve (adoption rate, derivative of logistic)
bell_curve <- dnorm(time, mean = 10, sd = 2.5) * 100

# Combine into dataframe
df <- data.frame(time, s_curve, bell_curve)

# Define cutoffs for adopter categories
mean_time <- 10
sd_time <- 2.5
cuts <- data.frame(
  x = c(mean_time - 2*sd_time, mean_time - sd_time, mean_time,
        mean_time + sd_time, mean_time + 2*sd_time),
  label = c("Innovators", "Early Adopters", "Early Majority", 
            "Late Majority", "Laggards"),
  col = c("red", "orange", "green3", "darkgreen", "blue")
)

# Plot
ggplot(df, aes(x = time)) +
  geom_line(aes(y = s_curve, colour = "S-curve (Cumulative Adoption)"), size = 1) +
  geom_line(aes(y = bell_curve, colour = "Bell Curve (Adoption Rate)"), size = 1, linetype = "dashed") +
  geom_vline(data = cuts, aes(xintercept = x, colour = label), linetype = "dashed") +
  scale_colour_manual(
    name = "Legend",
    values = c("S-curve (Cumulative Adoption)" = "purple",
               "Bell Curve (Adoption Rate)" = "deeppink",
               "Innovators" = "red",
               "Early Adopters" = "orange",
               "Early Majority" = "goldenrod3",
               "Late Majority" = "darkgreen",
               "Laggards" = "blue")
  ) +
  labs(
    title = "S-curve and Bell Curve",
    x = "Time step",
    y = "Adoption percentage"
  ) +
  theme_minimal(base_size = 14)
